---
title: "Analysis on Indonesian 10y Sovereign Bonds Yield"
author: "Dea Avega Editya"
date: "8/5/2021"
output: html_document
---

#Outlines:

1. Introduction:

a. Literature review:
- Description of Local Currency Bonds 10y, its roles and yield,
- Description of Determinant factors:
US Treasury 10y, Credit Default Swap 5y & 10y, foreign ownership, policy rate (Bank Indonesia 7-day repo rate), exchange rate (Jakarta Interbank Offered Rate/JIBOR), Issuance Mechanism in Primary Market (new policy of greenshoe auction in 2020)

b. Motivation: 
- Previous researches are mixed on what determining bonds yield and only few of them cover Indonesia,
- Understanding factors that determining the 10y-bonds yield in Indonesia,
  


c. Data sources + available links

d. Scope and Limitation
Cant see average bid from each group (i.e. conventional banks, foreign, insurance/pension, others)

2. Methodology

Many confounding factors, hence visualizing with plots maybe can explain  
Understanding the Relationship through EDA

a. UST vs Yield

b. CDS vs Yield

c. Foreign Ownership vs Yield

d. Bank Indo Rate vs Yield

e. Exchange rate (JIBOR) vs Yield

f. Greenshoe mechanism vs Yield (also need to check historical Weighted Average Yield-awarded before and after greenshoe)

3. Key Findings

Unknown yet
i.e.: 

any changed patterns?

factors that steadily affect the yield?

pandemic effect? i.e. holding spending for investing in safe instrument? (need to check ownership of banks, mutual funds, insurance) 

shifting power (bonds ownership) foreign to domestic participant (Do central bank/domestic banks become more dominant)?

suplementary auction (greenshoe) can reduce yield in 2020-2021?


Summary

Bibliography


#Timeline

Chapter 1 + Data collection/conversion from excel tables 31 Aug 2021,

Chapter 2 1 Sep - 25 Oct 2021,

Chapter 3 26 Oct - 12 Nov 2021,

Summary, Bibliography, cleaning Code, creating slides 13 - 19 Nov 2021


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(tidyverse)
library(readr)
library(readxl)
library(stringr)
library(lubridate)
library(janitor)
library(cleaner)
library(plotly)
library(tsibble)
library(openxlsx)
library(zoo)
library(purrr)
library(sugrrants)
library(GGally)

#script for create UST, Yield, cds, foreign
fac1 <- readxl::read_xlsx("data/forecast_new.xlsx", col_names = T, sheet = 2) %>%
  clean_names() %>% as_tibble()

#read inflation
cpi <- read_csv("data/cpi_worldbank.csv", skip = 4) %>% 
  filter(`Country Name` == "Indonesia") %>% as_tibble()

cpi_t <- t(cpi[-1:-4]) 
colnames(cpi_t) <- cpi[ ,1]

cpi_t <- cbind(year = rownames(cpi_t), cpi_t)
rownames(cpi_t) <- NULL

cpi_t <- cpi_t %>% as_tibble() %>% 
  mutate(year = as.numeric(year), inflation = as.numeric(Indonesia)) %>% 
  select(-2) %>% drop_na()

#read ownership
own <- readxl::read_xlsx("data/kepemilikan asing.xlsx", col_names = T) %>% drop_na() %>%  janitor::clean_names() %>%  as_tibble()

own_total <- read_xlsx("data/TOTAL KEPEMILIKAN.xlsx", col_names = F) %>%
  rename(date = 1, ownership = 2) %>% 
  as_tibble()

ownership <- left_join(x = own, y = own_total, by = "date") %>% 
  filter(date >= date("2015-01-02"))

#combine all factors
yield_join <- left_join(fac1, ownership, by = "date") %>% 
  fill(c(ownership, foreign_own), .direction = "down") %>% 
  mutate(foreign_pct = foreign_own/ownership, year = year(date))

yield_factor <- left_join(yield_join, cpi_t, by = "year") %>% select(-year) %>% 
  na_replace(inflation, replacement = 1.4)

#impute inflation for 2021
na_replace(yield_factor$inflation, replacement = 1.4)

yield_factor <- yield_factor %>% 
  mutate(for_chg = foreign_pct - lag(foreign_pct, default = first(foreign_pct)), 
         up_down = case_when(for_chg < 0 ~ "down",
                             for_chg > 0 ~ "up", TRUE ~ "no change")) %>% 
  rename(domestic_10y = domestik_10y)

#add policy rate
pol_rate <- read_xlsx("data/policy_rate.xlsx", skip = 4, col_names = T) %>% 
  mutate(date = as.Date(Tanggal), rate = as.double(parse_number(`BI-7Day-RR`))) %>%
  select(date, rate)

yield_factors <- left_join(yield_factor, pol_rate, by = "date") %>% 
  fill(rate, .direction = "down")

unique(ownership_bonds$group)

cek <- ownership_bonds %>% 
  filter()
  summarise(date, id, )

# yield_factors <- left_join(yield_factor)

```

```{r}
yield_factor %>% 
  mutate(year = year(date)) %>% 
  group_by(year) %>% 
  summarise(avg_10y = mean(domestik_10y), inflation) %>% 
  ggplot(aes(x = avg_10y, y = inflation)) + 
  geom_point()

#foreign changes
yield_factor %>% 
  mutate(week_year = yearweek(date)) %>% 
  group_by(week_year) %>% 
  summarise(avg_10y = mean(domestic_10y), avg_chg = mean(foreign_pct)) %>% 
  ggplot(aes(x = avg_chg, y = avg_10y)) + 
  geom_point()

```

Several studies indicates the relationship between bonds yield and foreign ownership (i.e......). We will observe the relationship between these two variables through visualization approach. 

By just putting the number, we can not see any clear pattern from these variables. However, if we facet the plot by year, now obvious patterns of the relationship arises.

```{r foreign_yield, fig.cap = "Foreign Ownership and Domestic 10y Yield"}
ownership_2015 %>% 
  mutate(cg = factor(case_when(group == "government" ~ "bi", TRUE ~ "non"))) %>% 
  ggplot(aes(x = id, y = value/1000, fill = cg))+
  geom_col()+
  coord_flip()+
  theme(legend.title = element_blank()) 


yield_factor %>% 
  mutate(year = year(date)) %>% 
  ggplot(aes(x = foreign_pct, y = domestic_10y)) + 
  geom_point() +
  facet_wrap(~ year)

yield_factors %>% 
  mutate(date= date(date), year = year(date), bi_pct = central_bank/tot_ownership, con_pct = con_bank/tot_ownership, islam_pct = islam_bank/tot_ownership, ind_pct = individual/tot_ownership) %>% 
  ggplot(aes(y = domestic_10y)) + 
  geom_point(aes(x = foreign_pct)) + 
  facet_wrap(~ year)

## insurance and pension fund strongly positive correlation
yield_factors %>% 
  mutate(date= date(date), year = year(date), bi_pct = central_bank/tot_ownership, con_pct = con_bank/tot_ownership, islam_pct = islam_bank/tot_ownership, ind_pct = individual/tot_ownership, ins_pct = ins_pension/tot_ownership) %>% 
  ggplot(aes(y = domestic_10y)) + 
  geom_point(aes(x = ins_pct)) + 
  facet_wrap(~ year)

yield_factors %>% 
  mutate(date= date(date), year = year(date), bi_pct = central_bank/tot_ownership, con_pct = con_bank/tot_ownership, islam_pct = islam_bank/tot_ownership, ind_pct = individual/tot_ownership, ins_pct = ins_pension/tot_ownership, other_pct = others/tot_ownership) %>% 
  ggplot(aes(y = domestic_10y)) + 
  geom_point(aes(x = other_pct))+
  facet_wrap(~ year)

yield_factors %>% 
  mutate(date= date(date), year = year(date), bi_pct = central_bank/tot_ownership, con_pct = con_bank/tot_ownership, islam_pct = islam_bank/tot_ownership, ind_pct = individual/tot_ownership, ins_pct = ins_pension/tot_ownership, other_pct = others/tot_ownership) %>% 
  ggplot(aes(y = domestic_10y)) + 
  geom_point(aes(x = ind_pct))+
  facet_wrap(~ year)

library(GGally)

yield_factor %>% 
  filter(year(date) == 2016) %>% 
GGally::ggscatmat()

yield_factor %>% 
  filter(year(date) == 2018) %>% 
GGally::ggscatmat()

yield_factor %>% 
GGally::ggscatmat(columns = c(2,3,4,5))

yield_factors <- read_rds("yield_factors.rds")

###add auction days
yield_factors <- yield_factors %>% 
  mutate(auction_day = case_when(
    date(date) %in% auction_date$value ~ "yes",
                TRUE ~ "no"))

yield_factors <- yield_factors %>% 
  mutate(auction_day = case_when(auction_day == "no" & year == 2015 ~ "", TRUE ~ auction_day))

inds <- which(yield_factors$auction_day == "yes")

yield_factors[inds-3, "pre_auction"] <- yield_factors[inds, "auction_day"]

yield_factors[inds-2, "pre_auction"] <- yield_factors[inds, "auction_day"]

yield_factors[inds-1, "pre_auction"] <- yield_factors[inds, "auction_day"]

yield_factors[inds-0, "pre_auction"] <- yield_factors[inds, "auction_day"]

inds_yes <- which(yield_factors$pre_auction == "yes")

yield_factors[inds_yes, "auction_day"] <- yield_factors[inds_yes, "pre_auction"]

### auction vs non_auc days
average_2017 <- yield_factors %>% filter(year == 2017) %>% 
  mutate(month = month(date)) %>% 
  group_by(month) %>%  summarise(mean = mean(domestic_10y))

average_2018 <- yield_factors %>% filter(year == 2018) %>% 
  mutate(month = month(date)) %>% 
  group_by(month) %>%  summarise(mean = mean(domestic_10y))

average_2019 <- yield_factors %>% filter(year == 2019) %>% 
  mutate(month = month(date)) %>% 
  group_by(month) %>%  summarise(mean = mean(domestic_10y))

average_2020 <- yield_factors %>% filter(year == 2020) %>% 
  mutate(month = month(date)) %>% 
  group_by(month) %>%  summarise(mean = mean(domestic_10y))

average_2021 <- yield_factors %>% filter(year == 2021) %>% 
  mutate(month = month(date)) %>% 
  group_by(month) %>%  summarise(mean = mean(domestic_10y))
  
yield_factors %>% filter(year == 2017) %>% 
  mutate(month = month(date)) %>% 
  group_by(month) %>%   
ggplot(aes(x = factor(day(date)), y = domestic_10y, color = auction_day, alpha = 0.1))+
  geom_point()+scale_x_discrete(breaks = seq(1, 31, 4))+
  geom_hline(data = average_2017, aes(yintercept = mean), alpha = 0.5)+
  facet_wrap(~month)

yield_factors %>% filter(year == 2018) %>% 
  mutate(month = month(date)) %>% 
  group_by(month) %>%   
ggplot(aes(x = factor(day(date)), y = domestic_10y, color = auction_day, alpha = 0.1))+
  geom_point()+scale_x_discrete(breaks = seq(1, 31, 4))+
  geom_hline(data = average_2018, aes(yintercept = mean), alpha = 0.5)+
  facet_wrap(~month)

yield_factors %>% filter(year == 2019) %>% 
  mutate(month = month(date)) %>% 
  group_by(month) %>%   
ggplot(aes(x = factor(day(date)), y = domestic_10y, color = auction_day, alpha = 0.1))+
  geom_point()+scale_x_discrete(breaks = seq(1, 31, 4))+
  geom_hline(data = average_2019, aes(yintercept = mean), alpha = 0.5)+
  facet_wrap(~month)

yield_factors %>% filter(year == 2020) %>% 
  mutate(month = month(date)) %>% 
  group_by(month) %>%   
ggplot(aes(x = factor(day(date)), y = domestic_10y, color = auction_day, alpha = 0.1))+
  geom_point()+scale_x_discrete(breaks = seq(1, 31, 4))+
  geom_hline(data = average_2020, aes(yintercept = mean), alpha = 0.5)+
  facet_wrap(~month)

yield_factors %>% filter(year == 2021) %>% 
  mutate(month = month(date)) %>% 
  group_by(month) %>%   
ggplot(aes(x = factor(day(date)), y = domestic_10y, color = auction_day, alpha = 0.1))+
  geom_point()+scale_x_discrete(breaks = seq(1, 31, 4))+
  geom_hline(data = average_2021, aes(yintercept = mean), alpha = 0.5)+
  facet_wrap(~month)

cek <- yield_factors %>% 
  group_by(auction_day) %>% select(!pre_auction)
  # mutate(domestic_10y = as.numeric(domestic_10y)) %>% 

cek %>% filter(year == 2016) %>% 
  mutate(month = month(date)) %>% 
  ggplot(aes(y = domestic_10y, color = auction_day)) +
  geom_boxplot()+facet_wrap(~month)

cek %>% filter(year == 2017) %>% 
  mutate(month = month(date)) %>% 
  ggplot(aes(y = domestic_10y, color = auction_day)) +
  geom_boxplot()+facet_wrap(~month)
  
cek %>% filter(year == 2018) %>% 
  mutate(month = month(date), date = date(date)) %>% 
  ggplot(aes(y = domestic_10y, x= factor(date), color = auction_day)) +
  geom_point()+facet_wrap(~month)

cek %>% filter(year == 2020) %>% 
  mutate(month = month(date)) %>% 
  ggplot(aes(y = domestic_10y, color = auction_day)) +
  geom_boxplot()+facet_wrap(~month)+theme(legend.position = "bottom")


cek %>% filter(year == 2018) %>% 
  mutate(month = month(date)) %>% 
  ggplot(aes(x = domestic_10y, color = auction_day), group = auction_day) +
  geom_density()+
               facet_wrap(~ month)







# yes <- wider$yes %>% na.omit()
# no <- wider$no %>% na.omit()
# 
# yes_no <- cbind(yes, no) %>% as_tibble()
# 
# k2 <- kmeans(yes_no, center = 2, nstart = 25)
# 
# fviz_cluster(k2, data = yes_no)

model <- lm(domestic_10y ~ cds_5y + cds_10y + ust_10y + foreign_pct + rate, 
            data = yield_factors %>% select(!date, !for_chg, !up_down, !tot_ownership, !foreign, !central_bank, !individual, !ins_pension, !mutual_fund, !securities_company, !con_bank, !islam_bank, !others, !year))

model19 <- lm(domestic_10y ~ cds_5y + cds_10y + ust_10y + foreign_pct, data = yield_factor %>% mutate(year = year(date)) %>%  filter(year == 2019))

model20 <- lm(domestic_10y ~ cds_5y + cds_10y + ust_10y + foreign_pct, data = yield_factor %>% mutate(year = year(date)) %>%  filter(year == 2020))

# reduce to "best fit" model with

summary(model)
model_best <- step(model, trace = FALSE)
summary(model_best)

#diagnose model
ggnostic(model_best)

```

As we can see from the plot, in 2015 and 2016 foreign ownership seems to have no strong relationship with the yield. Different situation happened in 2017-2019 where the changes in foreign ownership seems to strong-negatively affect the bonds yield movement. 

To mention, 2017 is the year when Indonesia got investment grade rating from S&P, following FITCH and Moody's in previous years. It means that broader category of foreign investment entities (i.e. pension funds and insurance) can enter the country's market since Indonesia's rating has fulfilled their criteria of investment (yunianto, 2018). 

Particularly In 2020, the pattern is quite anomaly in which increase from 0.25-0.33% in foreign ownership seems to raise the yield from 6 up to 8.5%. The yield stumbles afterwards with the increase of foreign ownership up to 40%. 

In January-August 2021, the relationship looks non-linear with no obvious pattern. The ownership drops below 25% but interestingly yield decreases further (6-7%).

We may assume that the anomaly in 2020 and 2021 are due to Covid19 pandemic that occurred since early 2020. The government is increasing its funding to sustain the economy. The decrease in foreign portion could be because foreigners sell their bonds more than their buying (net sell), or another reason is because their portion is deluged by domestic participants (i.e. central bank), thus we will track the pattern of domestic ownership in these years to check our assumption.

```{r}
#load bonds ownership
ownership_bonds <- readRDS('ownership_bonds.rds') %>% 
  mutate(value = as.double(value), type = factor(type), month = month(date), year = year(date), id = trimws(gsub(".*-", "", id)))

ownership_bonds %>% 
filter(date == "2021-08-02", group %in% c("government (part of BI)", "government"))


ownership_bonds <- ownership_bonds %>% 
  mutate(id = case_when(id %in% c("insurance company", "insurance", "pension funds", "insurance and pension fund", "pension fund") ~ "insurance & pension fund", 
                        id %in% c("mutual fund", "mutual funds") ~ "mutual fund",
                        TRUE ~ id)) %>% 
  group_by(id, date, group, type) %>% summarise(id, date, month, year, group, type, value = sum(value)) %>% 
  distinct()

unique(ownership_bonds$group)

ownership_bonds %>% 
filter(group == "government", date == "2015-12-31")

ownership_bonds %>%
  filter(group %in% c("bank", "government", "non-bank"), year != 2015) %>% 
  group_by(id, month, year, group) %>%
  ggplot(aes(x = factor(month), y = value,  fill = id))+
    geom_col(width = 5)+
facet_wrap(~ year, ncol = 2)
  # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

by_id <- ownership_bonds %>% 
  filter(type == "total") %>% 
  summarise(id, date, value) 


total_bonds <- by_id %>% 
  group_by(date) %>% 
  filter(group %in% c("government", "bank", "non-bank")) %>% 
  summarize(date, value = sum(value)) %>% 
  distinct() %>% 
  rename(tot_ownership = value)

ownership_bonds %>% 
  group_by(id) %>% 
  filter(date == "2015-12-31", group != "non-bank (part of foreign holders)")

#add total ownership
yield_factors <- left_join(yield_factor, total_bonds, by = "date") %>% 
  fill(tot_ownership, .direction = "down")

foreign <-by_id %>% 
  group_by(date) %>% 
  filter(id == "foreign holders") %>% 
  summarize(date, value = sum(value)) %>% 
  distinct() %>% 
  rename(foreign = value)

yield_factors <- left_join(yield_factors, foreign, by = "date") %>% 
  fill(foreign, .direction = "down")

yield_factors <- yield_factors %>% 
  select(-foreign_own, - ownership)

central_bank <- by_id %>% 
  group_by(date) %>% 
  filter(id == "bank indonesia") %>% 
  summarize(date, value = sum(value)) %>% 
  distinct() %>% 
  rename(central_bank = value)

yield_factors <- left_join(yield_factors, central_bank, by = "date") %>% 
  fill(central_bank, .direction = "down")

individual <- by_id %>% 
  group_by(date) %>% 
  filter(id == "individual") %>% 
  summarize(date, value = sum(value)) %>% 
  distinct() %>% 
  rename(individual = value)

yield_factors <- left_join(yield_factors, individual, by = "date") %>% 
  fill(individual, .direction = "down")

ins_pension <- by_id %>% 
  group_by(date) %>% 
  filter(id == "insurance & pension fund") %>% 
  summarize(date, value = sum(value)) %>% 
  distinct() %>% 
  rename(ins_pension = value)

yield_factors <- left_join(yield_factors, ins_pension, by = "date") %>% 
  fill(ins_pension, .direction = "down")


mutual <- by_id %>% 
  group_by(date) %>% 
  filter(id == "mutual fund") %>% 
  summarize(date, value = sum(value)) %>% 
  distinct() %>% 
  rename(mutual_fund = value)

yield_factors <- left_join(yield_factors, mutual, by = "date") %>% 
  fill(mutual_fund, .direction = "down")

securities <- by_id %>% 
  group_by(date) %>% 
  filter(id == "securities company") %>% 
  summarize(date, value = sum(value)) %>% 
  distinct() %>% 
  rename(securities_company = value)

yield_factors <- left_join(yield_factors, securities, by = "date") %>% 
  fill(securities_company, .direction = "down")

unique(by_id$id)

con_bank <- by_id %>% 
  group_by(date) %>% 
  filter(id == "conventional bank") %>% 
  summarize(date, value = sum(value)) %>% 
  distinct() %>% 
  rename(con_bank = value)

yield_factors <- left_join(yield_factors, con_bank, by = "date") %>% 
  fill(con_bank, .direction = "down")

islam_bank <- by_id %>% 
  group_by(date) %>% 
  filter(id == "islamic bank") %>% 
  summarize(date, value = sum(value)) %>% 
  distinct() %>% 
  rename(islam_bank = value)

yield_factors <- left_join(yield_factors, islam_bank, by = "date") %>% 
  fill(islam_bank, .direction = "down")

others <- by_id %>% 
  group_by(date) %>% 
  filter(id == "others") %>% 
  summarize(date, value = sum(value)) %>% 
  distinct() %>% 
  rename(others = value)

yield_factors <- left_join(yield_factors, others, by = "date") %>% 
  fill(others, .direction = "down")

yield_factors <- yield_factors %>% 
  mutate(year = year(as.Date(date, tz = "UTC")))

pol_rate <- read_xlsx("data/policy_rate.xlsx", skip = 4, col_names = T) %>% 
  mutate(date = as.Date(Tanggal), rate = as.double(parse_number(`BI-7Day-RR`))) %>%
  select(date, rate)

yield_factors <- left_join(yield_factors, pol_rate, by = "date") %>% 
  fill(rate, .direction = "down")

saveRDS(yield_factors, "yield_factors.rds")

#scattermatrix
colnames(yield_factors)

yield_factors %>% 
  filter(year > 2019) %>% 
GGally::ggscatmat(columns = c(2,3,4,5))

unique(colnames(yield_factors))

# yield_factors %>% 
#   filter(year > 2019) %>% 
#   mutate(bi_pct = central_bank/tot_ownership, indv_pct = individual/tot_ownership, insp_pct = ins_pension/tot_ownership, mut_pct = mutual_fund/tot_ownership, sec_pct = securities_company/tot_ownership, conv_pct = con_bank/tot_ownership, isl_pct = islam_bank/tot_ownership) %>% 
#   select(domestic_10y, foreign_pct, bi_pct, indv_pct, insp_pct, mut_pct, sec_pct, conv_pct, isl_pct) %>% 
# GGally::ggscatmat()


yield_factors %>% 
  filter(year > 2018) %>% 
GGally::ggscatmat(columns = c(5, 13, 14, 15, 16, 17, 18, 19))


model <- lm(domestic_10y ~ cds_5y + ust_10y + foreign + rate + central_bank, data = yield_factors)

# reduce to "best fit" model with
model_best <- step(model, trace = FALSE)
summary(model_best)

#diagnose model
ggnostic(model_best)

```

As shown in the plot, proportion of Bank Indonesia and Conventional Bank are getting bigger in 2020-2021, while proportion of foreign holders is diluted. This is can be seen as a result of implementation of new regulation on 1 May 2020 that mandate banks to reserve government bonds. This regulation has an impact on reducing bonds yield further in these two consecutive years.

Another noticeable change is proportion of individual in the bonds ownership that is growing started from mid of 2020 up to 2021. We may argue that people who prefer to hold their spending is channelling their excess money to investment, especially in a safe instrument like government bonds.

## Policy Rate
```{r}

yield_factors %>% 
  na.omit() %>% 
  mutate(year = year(date)) %>% 
   filter(year != 2015) %>% 
  select(date, domestic_10y, rate, foreign, central_bank) %>%
  gather(key = "variable", value = "value", -date) %>% 
  ggplot(aes(x = date, y = log(value)))+
  geom_smooth(aes(color = variable))+theme_minimal()

## ownership pct
yield_factors %>% 
  na.omit() %>% 
  mutate(date= date(date), year = year(date), bi_pct = central_bank/tot_ownership, con_pct = con_bank/tot_ownership, islam_pct = islam_bank/tot_ownership, ind_pct = individual/tot_ownership, other_pct = others/tot_ownership) %>% 
   filter(year != 2015) %>% 
  select(date, foreign_pct, bi_pct, con_pct, islam_pct, ind_pct, other_pct) %>%
  gather(key = "variable", value = "value", -date) %>% 
  ggplot(aes(x = date, y = value))+
  geom_area(aes(fill = variable), alpha = 0.7)+
  theme_classic()
####Others = securities, corporations, and foundations.



yield_factors %>% na.omit() %>% 
  mutate(spread_ust = rate-ust_10y) %>% 
  ggplot(aes(x = date(date), y = ust_10y))+
  geom_smooth(color = "green")+geom_smooth(aes(y= inflation))+geom_smooth(aes(y=rate), color = "red")


# spread pol-rate vs yield
yield_factors %>% 
  na.omit() %>% 
  mutate(year = year(date), spread = abs(domestic_10y-rate)) %>%
  filter(year != 2015) %>% 
  select(date, year, spread) %>% 
  ggplot(aes(y = spread))+
  geom_boxplot()+
  facet_wrap(~year)

yield_factors %>% 
  filter(year > 2019)
  

```

From plot above, the movement of domestic 10y yield is parallel with movement of policy rate. The median spread in 2016 until 2018 is quite the same (around 2.5) while median spread of 2019 is the lowest. In 2020 and 2021, the differents between 10y yield and policy rate are increase with median spread is about 3. The range of spread in 2020 is also the widest compared to other observed years.



