---
title: "reserved"
author: "Dea Avega Editya"
date: "10/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r eval = FALSE}
library(broom)

# model <- lm(formula = domestic_10y ~ cds_5y + ust_10y + rate:year + foreign + ins_pension + foreign:year + ins_pension:year + auction_day:month,
#  data = data_model)

model <- lm(formula = domestic_10y ~ cds_5y + ust_10y + rate + foreign_pct + I(vix^2) + exchange_mid + auction_day,
            data = data_model)

summary(model)
tidy(model)

# highest vix in early pandemic 2020-03-16 
# MArch 2 2020 is the first case of covid
data_model %>% filter (vix == 82.69)


# best <- drop1(model, test = "F")
# summary(best)


aug <- broom::augment(model, data = data_model %>% filter(year != 2015) %>% drop_na())

aug %>% 
  ggplot(aes(x = .fitted, y = .resid))+
  geom_point()

aug %>%
  ggplot(aes(x = .cooksd, y = .hat))+
  geom_point()

# aug %>% 
#   ggplot(aes(x = .cooksd, y = .hat))+
#   geom_point()

outlier <- aug %>% filter(.cooksd > 0.05)

#not in
`%notin%` <- Negate(`%in%`)

data_model %>% filter(date %in% outlier$date)

####remove extreme outliers
data_model2 <- data_model %>% filter(date %notin% outlier$date)

# model <-  lm(formula = domestic_10y ~ cds_5y + ust_10y + rate + foreign:year + I(vix^2) + exchange_mid + auction_day:month, data = data_model2)


# model <- lm(formula = domestic_10y ~ cds_5y + ust_10y + rate + foreign:year + I(vix^2) + cpi_inflation + exchange_mid + auction_day:year,
#             data = data_model)

summary(model)
tidy(model)

aug <- broom::augment(model, data = data_model2 %>% filter(year != 2015) %>% drop_na())

aug %>% 
  ggplot(aes(x = .fitted, y = .resid))+
  geom_point()

aug %>% 
  ggplot(aes(x = .cooksd, y = .hat))+
  geom_point()

#check multicol
#explain GVIF variable 
car::vif(model)

# reduce to "best fit" model with

summary(model)
# model_best <- step(model, trace = FALSE)
# summary(model_best)

#diagnose model
ggnostic(model_best)

```


In particular, the results suggest that
improved macroeconomic fundamentals, such as higher net foreign assets (in
terms of GDP or imports), lower fiscal deficits, and lower ratios of debt service to
exports and debt to GDP, help to lower sovereign spreads @arora

a. UST vs Yield
While the dramatic rise in capital flows to
emerging markets has been induced primarily by the implementation of sound
macroeconomic policies and wide structural reforms in these countries, it has
also been driven by changing conditions in industrial countries that have
encouraged investors to diversify their portfolios into developing country
assets.  Interest rate spreads (the differences between yields on
sovereign bonds of developing countries and U.S. treasury securities of
comparable maturities), which are a proxy for country risk, have tended to
move in the same direction as the changes in U.S. interest rates @arora

b. CDS vs Yield

Sovereign CDS spreads are used as an indicator of foreign currency sovereign
creditworthiness. Lower sovereign CDS spreads are expected to lower local currency sovereign bond yields @gadanecz

c. Foreign Ownership vs Yield

d. Bank Indo Rate vs Yield

e. Exchange rate (JISDOR) vs Yield
Investors are exposed to gains and losses from exchange rate movements on their holdings of local currency sovereign bond. exchange rate risk can represent an important channel of transmission of market sentiment, uncertainty and default risk to local currency bond yield 
Exchange rate risk tends to affect liquidity
conditions in both foreign exchange and domestic bond markets, which tend to be relatively low in many EMEs even in tranquil times.
The direction of the
causality runs from exchange rate volatility to local currency sovereign bond yields. This is especially the case in Asia and eastern
Europe. In these two regions, local currency sovereign bond markets are relatively liquid and foreign participation relatively large.

The sensitivity of EME local currency sovereign bond yields to exchange rate volatility increases after
the global financial crisis, and further after the taper tantrum in mid-2013
@gadanecz


f. Volatility index (VIX) vs Yield
It is a measure of market expectations of near-term volatility conveyed by S&P500 stock index option prices and considered as a forward-looking measure of investor risk. Hartelius et al. (2008) highlights the strong dependence of emerging market returns to the VIX, which should be positively
related to changes in emerging market spreads since more risk aversion increases spreads. An attractive feature of this index is that it can
be considered as exogenous for emerging economies (Siklos, 2011). @hajer

This result could be explained by the fact that as investors become more risk-averse and seek safer assets, the expected growth in volatility encourages them to liquidate their positions in risky assets in favor of safer ones, thus increasing sovereign spreads.
@hajer

Intended to capture changes in investor sentiment which may be related to
expected changes in U.S. monetary policy. It may also pick up the effects of other market-related events, such as the flight to quality effects during the Asian crisis.
@arora

As historical data demonstrates a strong negative correlation of volatility to the stock market returns – that is, when stock returns go down, volatility rises and vice versa.(investopedia)



# Conclusion






factors that steadily affect the yield? (foreign and insurance/pension fund are steadily associated with yield movement. Foreign has strongly negative correlation until 2019, while insurance/pension has strongly positive correlation)

pandemic effect? i.e. holding spending for investing in safe instrument? (need to check ownership of domestic banks, mutual funds, insurance, individual investors) 

shifting power (bonds ownership) foreign to domestic participant (Do central bank/domestic banks become more dominant)?
(Quantitative easing of Bank Indonesia and mandatory purchase of domestic banks can push down the yield)


```{r }

# 
# 
# #check debt_gdp ratio vs yield
# yield_factors %>% mutate(year = year(date)) %>% 
#   ggplot(aes(y = domestic_10y)) + 
#   geom_jitter(aes(x = debt_gdp))+
#   facet_wrap(~year)

```


```{r foreign_yield, fig.cap = "Foreign Ownership and Domestic 10y Yield", eval=FALSE}

# yield_factors %>%  mutate(year = year(date), con_pct = con_bank/tot_ownership, bi_pct = central_bank/tot_ownership, ins_pct = ins_pension/tot_ownership) %>% 
#   # filter(year == 2019) %>% 
#   select(date, foreign, central_bank, individual, mutual_fund, con_bank, islam_bank, ins_pension, others) %>%
#   gather(key = "variable", value = "value", -date) %>% 
#   ggplot(aes(x = variable, y = value/1000, fill = value))+
#   geom_col()+
#   coord_flip()+
#   theme(legend.title = element_blank())


# yield_factors %>% 
#   mutate(year = year(date), con_pct = con_bank/tot_ownership, bi_pct = central_bank/tot_ownership, ins_pct = ins_pension/tot_ownership) %>% 
#   ggplot(aes(y = domestic_10y))+
#   geom_point(aes(x = con_pct, color = "red"))+
#   geom_point(aes(x = bi_pct))+
#   theme(legend.title = element_blank())+facet_wrap(~ year)

## others has strong positive correlation with yield
yield_factors %>% 
mutate(year = year(date), mon = month(date), con_pct = con_bank/tot_ownership, bi_pct = central_bank/tot_ownership, others_pct = others/tot_ownership) %>% 
  ggplot(aes(x = others, y = domestic_10y)) + 
  geom_point() +
  facet_wrap(~ year)

#check ownership in auction days
# yield_factors %>%
# mutate(year = year(date), month = month(date), con_pct = scale(con_bank/tot_ownership), bi_pct = central_bank/tot_ownership, others_pct = scale(others/tot_ownership), ins_pct = scale(ins_pension/tot_ownership)) %>%
#   filter(year == 2019, month == 2) %>%
#   ggplot(aes(x = date)) +
#   geom_line(aes(y = ins_pct))+
#   geom_line(aes(y = con_pct, color = "red"))+
#   geom_point(aes(y = auction_day, color = factor(auction_day)))


# bi and con_bank is balancing each other
# since 2020, they move in the same direction
# burden sharing for pandemic
yield_factors %>% 
mutate(year = year(date), con_pct = con_bank/tot_ownership, bi_pct = central_bank/tot_ownership, others_pct = scale(others/tot_ownership), ins_pct = ins_pension/tot_ownership) %>% 
  ggplot(aes(x = date)) + 
  geom_line(aes(y = bi_pct))+
  geom_line(aes(y = con_pct))

###CDS vs UST spread
# yield_factors %>% 
#   mutate(year = year(date)) %>% 
#   mutate(date= date(date), month =  month(date), spread_10 = cds_10y-domestic_10y, spread_5 = cds_5y-domestic_10y, spread_ust = domestic_10y-ust_10y) %>%  
#   filter(year == 2021) %>% 
#   ggplot(aes(x = date)) + 
#   geom_line(aes(y = sqrt(spread_10)))+
#   geom_line(aes(y = sqrt(spread_5)), color = "red", alpha = 0.5)+
#   geom_line(aes(y = spread_ust), color = "blue")+
#    scale_x_date(labels = date_format("%m-%Y"))


### CDS vs centralbank rate spread

# yield_factors %>% 
#   mutate(year = year(date)) %>% 
#   mutate(date= date(date), month =  month(date), spread_10 = cds_10y-domestic_10y, spread_5 = cds_5y-domestic_10y, spread_bi = domestic_10y-rate) %>%  
#   filter(year == 2021) %>% 
#   ggplot(aes(x = date)) + 
#   geom_line(aes(y = sqrt(spread_10)))+
#   geom_line(aes(y = sqrt(spread_5)), color = "red", alpha = 0.5)+
#   geom_line(aes(y = spread_bi), color = "blue")+
#    scale_x_date(labels = date_format("%m-%Y"))


# ust vs policy rate vs spread rate-ust
# check how differ central bank rate from UST
# yield_factors %>% 
#   mutate(year = year(date)) %>% 
#   mutate(date= date(date), month =  month(date), spread_10 = cds_10y-domestic_10y, spread_5 = cds_5y-domestic_10y, spread_ust = domestic_10y-ust_10y, spread_bi = domestic_10y-rate, ust_bi = rate-ust_10y) %>% 
#   filter(year > 2015) %>% 
#   ggplot(aes(x = date)) + 
#   geom_area(aes(y = ust_bi), fill = "darkgreen", alpha = 0.5)+
#   geom_line(aes(y = rate))+
#   geom_line(aes(y = ust_10y), color = "red")+
#   # geom_line(aes(y = spread_ust))+
#   # geom_line(aes(y = spread_bi), color = "red", alpha = 0.5)+
#   scale_x_date(labels = date_format("%m-%Y"))
  
# yield_factors %>% 
#   mutate(year = year(date)) %>% 
#   mutate(date= date(date), month =  month(date), spread_10 = cds_10y-domestic_10y, spread_5 = cds_5y-domestic_10y, spread_ust = domestic_10y-ust_10y, spread_bi = domestic_10y-rate, ust_bi = rate-ust_10y) %>%  
#   filter(year > 2015) %>% 
#   ggplot(aes(x = spread_ust, y = domestic_10y))+
#   geom_jitter()

###spread ust vs yield in 2019 and 2021 is not strongly associated
# yield_factors %>% 
#   mutate(year = year(date)) %>% 
#   mutate(date= date(date), month =  month(date), spread_10 = cds_10y-domestic_10y, spread_5 = cds_5y-domestic_10y, spread_ust = domestic_10y-ust_10y, spread_bi = domestic_10y-rate, ust_bi = rate-ust_10y) %>%  
#   filter(year > 2015) %>% 
#   ggplot(aes(x = spread_ust, y = domestic_10y))+
#   geom_jitter()+ facet_wrap(~year)

###spread bi vs yield 
# yield_factors %>% 
#   mutate(year = year(date)) %>% 
#   mutate(date= date(date), month =  month(date), spread_10 = cds_10y-domestic_10y, spread_5 = cds_5y-domestic_10y, spread_ust = domestic_10y-ust_10y, spread_bi = domestic_10y-rate, ust_bi = rate-ust_10y) %>%  
#   filter(year > 2015) %>% 
#   ggplot(aes(x = spread_bi, y = domestic_10y))+
#   geom_jitter()+
#   facet_wrap(~year)



# yield_factors %>% 
#   mutate(year = year(date)) %>% 
#   mutate(date= date(date), month =  month(date), spread_10 = cds_10y-domestic_10y, spread_5 = cds_5y-domestic_10y, spread_ust = domestic_10y-ust_10y, spread_bi = domestic_10y-rate, ust_bi = rate-ust_10y) %>%  
#   filter(year > 2015) %>% 
#   ggplot(aes(x = cds_10y, y = domestic_10y))+
#   geom_jitter()+
#   facet_wrap(~year)


# spread movement
# spread_2016 <- yield_factors %>% 
#   mutate(year = year(date)) %>% 
#   mutate(date= date(date), month =  month(date), spread_10 = domestic_10y - cds_10y/100, spread_5 = domestic_10y - cds_5y/100, spread_ust = domestic_10y-ust_10y) %>% 
#   filter(year == 2016) %>% 
#   ggplot(aes(x = date)) + 
#   geom_line(aes(y = spread_10))+
#   geom_line(aes(y = spread_5), color = "red", alpha = 0.5)+
#   geom_line(aes(y = spread_ust), color = "blue")+
#    scale_x_date(labels = date_format("%m-%Y"))


# spread_2017 <- yield_factors %>% 
#   mutate(year = year(date)) %>% 
#   mutate(date= date(date), month =  month(date), spread_10 = domestic_10y - cds_10y/100, spread_5 = domestic_10y - cds_5y/100, spread_ust = domestic_10y-ust_10y) %>% 
#   filter(year == 2017) %>% 
#   ggplot(aes(x = date)) + 
#   geom_line(aes(y = spread_10))+
#   geom_line(aes(y = spread_5), color = "red", alpha = 0.5)+
#   geom_line(aes(y = spread_ust), color = "blue")+
#    scale_x_date(labels = date_format("%m-%Y"))
# 
# spread_2018 <- yield_factors %>% 
#   mutate(year = year(date)) %>% 
#   mutate(date= date(date), month =  month(date), spread_10 = domestic_10y - cds_10y/100, spread_5 = domestic_10y - cds_5y/100, spread_ust = domestic_10y-ust_10y) %>% 
#   filter(year == 2018) %>% 
#   ggplot(aes(x = date)) + 
#   geom_line(aes(y = spread_10))+
#   geom_line(aes(y = spread_5), color = "red", alpha = 0.5)+
#   geom_line(aes(y = spread_ust), color = "blue")+
#    scale_x_date(labels = date_format("%m-%Y"))
# 
# spread_2019 <- yield_factors %>% 
#   mutate(year = year(date)) %>% 
#   mutate(date= date(date), month =  month(date), spread_10 = domestic_10y - cds_10y/100, spread_5 = domestic_10y - cds_5y/100, spread_ust = domestic_10y-ust_10y) %>% 
#   filter(year == 2019) %>% 
#   ggplot(aes(x = date)) + 
#   geom_line(aes(y = spread_10))+
#   geom_line(aes(y = spread_5), color = "red", alpha = 0.5)+
#   geom_line(aes(y = spread_ust), color = "blue")+
#    scale_x_date(labels = date_format("%m-%Y"))
# 
# spread_2020 <- yield_factors %>% 
#   mutate(year = year(date)) %>% 
#   mutate(date= date(date), month =  month(date), spread_10 = domestic_10y - cds_10y/100, spread_5 = domestic_10y - cds_5y/100, spread_ust = domestic_10y-ust_10y) %>% 
#   filter(year == 2020) %>% 
#   ggplot(aes(x = date)) + 
#   geom_line(aes(y = spread_10))+
#   geom_line(aes(y = spread_5), color = "red", alpha = 0.5)+
#   geom_line(aes(y = spread_ust), color = "blue")+
#    scale_x_date(labels = date_format("%m-%Y"))
# 
# spread_2021 <- yield_factors %>% 
#   mutate(year = year(date)) %>% 
#   mutate(date= date(date), month =  month(date), spread_10 = domestic_10y - cds_10y/100, spread_5 = domestic_10y - cds_5y/100, spread_ust = domestic_10y-ust_10y) %>% 
#   filter(year == 2021) %>% 
#   ggplot(aes(x = date)) + 
#   geom_line(aes(y = spread_10))+
#   geom_line(aes(y = spread_5), color = "red", alpha = 0.5)+
#   geom_line(aes(y = spread_ust), color = "blue")+
#   scale_x_date(labels = date_format("%m-%Y"))
# 
# 
# plot_spread <- patchwork::wrap_plots(spread_2016, spread_2017, spread_2018, spread_2019, spread_2020, spread_2021, ncol = 3)
###spread mostly below cds, except in 2016 and 2020 after pandemic in march

yield_factors %>% 
  mutate(date= date(date), year = year(date), bi_pct = central_bank/tot_ownership, con_pct = con_bank/tot_ownership, islam_pct = islam_bank/tot_ownership, ind_pct = individual/tot_ownership, foreign_pct = foreign/tot_ownership) %>% 
  ggplot(aes(y = domestic_10y)) + 
  geom_point(aes(x = foreign_pct)) + 
  facet_wrap(~ year)

## insurance and pension fund strongly positive correlation
yield_factors %>% 
  mutate(date= date(date), year = year(date), bi_pct = central_bank/tot_ownership, con_pct = con_bank/tot_ownership, islam_pct = islam_bank/tot_ownership, ind_pct = individual/tot_ownership, ins_pct = ins_pension/tot_ownership) %>% 
  ggplot(aes(y = domestic_10y)) + 
  geom_point(aes(x = ins_pct)) + 
  facet_wrap(~ year)

yield_factors %>% 
  mutate(date= date(date), year = year(date), bi_pct = central_bank/tot_ownership, con_pct = con_bank/tot_ownership, islam_pct = islam_bank/tot_ownership, ind_pct = individual/tot_ownership, ins_pct = ins_pension/tot_ownership, other_pct = others/tot_ownership) %>% 
  ggplot(aes(y = domestic_10y)) + 
  geom_point(aes(x = other_pct))+
  facet_wrap(~ year)

yield_factors %>% 
  mutate(date= date(date), year = year(date), bi_pct = central_bank/tot_ownership, con_pct = con_bank/tot_ownership, islam_pct = islam_bank/tot_ownership, ind_pct = individual/tot_ownership, ins_pct = ins_pension/tot_ownership, other_pct = others/tot_ownership, mutual_pct = mutual_fund/tot_ownership) %>% 
  ggplot(aes(y = domestic_10y)) + 
  geom_point(aes(x = mutual_fund))+
  facet_wrap(~ year)

## check ggscatmat
# yield_factors %>% 
#   mutate(year = year(date)) %>% 
#   filter(year == 2020) %>% 
#     select(domestic_10y, cds_5y, ust_10y, central_bank, exchange_mid) %>%
# GGally::ggscatmat()



# 
# yield_factors %>% filter(year == 2018) %>% 
#   mutate(month = month(date)) %>% 
#   group_by(month) %>%   
# ggplot(aes(x = factor(day(date)), y = domestic_10y, color = factor(auction_day), alpha = 0.1))+
#   geom_point()+scale_x_discrete(breaks = seq(1, 31, 4))+
#   geom_hline(data = average_2018, aes(yintercept = mean), alpha = 0.5)+
#   facet_wrap(~month)
# 
# yield_factors %>% filter(year == 2019) %>% 
#   mutate(month = month(date)) %>% 
#   group_by(month) %>%   
# ggplot(aes(x = factor(day(date)), y = domestic_10y, color = factor(auction_day), alpha = 0.1))+
#   geom_point()+scale_x_discrete(breaks = seq(1, 31, 4))+
#   geom_hline(data = average_2019, aes(yintercept = mean), alpha = 0.5)+
#   facet_wrap(~month)
# 
# yield_factors %>% filter(year == 2020) %>% 
#   mutate(month = month(date)) %>% 
#   group_by(month) %>%   
# ggplot(aes(x = factor(day(date)), y = domestic_10y, color = factor(auction_day), alpha = 0.1))+
#   geom_point()+scale_x_discrete(breaks = seq(1, 31, 4))+
#   geom_hline(data = average_2020, aes(yintercept = mean), alpha = 0.5)+
#   facet_wrap(~month)
# 
# yield_factors %>% filter(year == 2021) %>% 
#   mutate(month = month(date)) %>% 
#   group_by(month) %>%   
# ggplot(aes(x = factor(day(date)), y = domestic_10y, color = auction_day, alpha = 0.1))+
#   geom_point()+scale_x_discrete(breaks = seq(1, 31, 4))+
#   geom_hline(data = average_2021, aes(yintercept = mean), alpha = 0.5)+
#   facet_wrap(~month)

# cek <- yield_factors %>% 
#   group_by(auction_day) %>% select(!pre_auction)
#   # mutate(domestic_10y = as.numeric(domestic_10y)) %>% 

# yield_factors %>% mutate(year = year(date)) %>% 
#   filter(year == 2016) %>% 
#   mutate(month = month(date)) %>% 
#   ggplot(aes(y = domestic_10y, color = auction_day)) +
#   geom_boxplot()+facet_wrap(~month)

# cek %>% filter(year == 2017) %>% 
#   mutate(month = month(date)) %>% 
#   ggplot(aes(y = domestic_10y, color = auction_day)) +
#   geom_boxplot()+facet_wrap(~month)
#   
# cek %>% filter(year == 2018) %>% 
#   mutate(month = month(date), date = date(date)) %>% 
#   ggplot(aes(y = domestic_10y, x= factor(date), color = auction_day)) +
#   geom_point()+facet_wrap(~month)
# 
# cek %>% filter(year == 2020) %>% 
#   mutate(month = month(date)) %>% 
#   ggplot(aes(y = domestic_10y, color = auction_day)) +
#   geom_boxplot()+facet_wrap(~month)+theme(legend.position = "bottom")



# yield distribution in each year
# 2018 yield move from low in early year to high in late period of year
# different direction compared to other years

# 2021 yield tend to be restrained in the middle
# yield_factors %>% mutate(year = year(date)) %>% filter(year == 2021) %>%  
#   mutate(month = month(date)) %>% 
#   ggplot(aes(x = domestic_10y, color = auction_day), group = auction_day) +
#   geom_density()+
#                facet_wrap(~ month)


```

As we can see from the plot, in 2015 and 2016 foreign ownership seems to have no strong relationship with the yield. Different situation happened in 2017-2019 where the changes in foreign ownership seems to strong-negatively affect the bonds yield movement. 

To mention, 2017 is the year when Indonesia got investment grade rating from S&P, following FITCH and Moody's in previous years. It means that broader category of foreign investment entities (i.e. pension funds and insurance) can enter the country's market since Indonesia's rating has fulfilled their criteria of investment (yunianto, 2018). 

Particularly In 2020, the pattern is quite anomaly in which increase from 0.25-0.33% in foreign ownership seems to raise the yield from 6 up to 8.5%. The yield stumbles afterwards with the increase of foreign ownership up to 40%. 

In January-August 2021, the relationship looks non-linear with no obvious pattern. The ownership drops below 25% but interestingly yield decreases further (6-7%).

We may assume that the anomaly in 2020 and 2021 are due to Covid19 pandemic that occurred since early 2020. The government is increasing its funding to sustain the economy. The decrease in foreign portion could be because foreigners sell their bonds more than their buying (net sell), or another reason is because their portion is deluged by domestic participants (i.e. central bank), thus we will track the pattern of domestic ownership in these years to check our assumption.

```{r}

#scattermatrix
# yield_factors %>% 
#   filter(year > 2019) %>% 
# GGally::ggscatmat(columns = c(2,3,4,5))

# yield_factors %>% 
#   filter(year > 2019) %>% 
#   mutate(bi_pct = central_bank/tot_ownership, indv_pct = individual/tot_ownership, insp_pct = ins_pension/tot_ownership, mut_pct = mutual_fund/tot_ownership, sec_pct = securities_company/tot_ownership, conv_pct = con_bank/tot_ownership, isl_pct = islam_bank/tot_ownership) %>% 
#   select(domestic_10y, foreign_pct, bi_pct, indv_pct, insp_pct, mut_pct, sec_pct, conv_pct, isl_pct) %>% 
# GGally::ggscatmat()

# yield_factors %>% 
#   filter(year > 2018) %>% 
# GGally::ggscatmat(columns = c(5, 13, 14, 15, 16, 17, 18, 19))

model <- lm(domestic_10y ~ cds_5y + ust_10y + foreign + rate + con_bank, data = yield_factors)

# reduce to "best fit" model with
model_best <- step(model, trace = FALSE)
summary(model)

#diagnose model
ggnostic(model)

```

As shown in the plot, proportion of Bank Indonesia and Conventional Bank are getting bigger in 2020-2021, while proportion of foreign holders is diluted. This is can be seen as a result of implementation of new regulation on 1 May 2020 that mandate banks to reserve government bonds. This regulation has an impact on reducing bonds yield further in these two consecutive years.

Another noticeable change is proportion of individual in the bonds ownership that is growing started from mid of 2020 up to 2021. We may argue that people who prefer to hold their spending is channelling their excess money to investment, especially in a safe instrument like government bonds.


Policy Rate


```{r eval = FALSE}

# yield_factors %>% 
#   na.omit() %>% 
#   mutate(year = year(date)) %>% 
#    filter(year != 2015) %>% 
#   select(date, domestic_10y, rate, foreign, central_bank) %>%
#   gather(key = "variable", value = "value", -date) %>% 
#   ggplot(aes(x = date, y = log(value)))+
#   geom_smooth(aes(color = variable))+theme_minimal()

## ownership pct
yield_factors %>% 
  na.omit() %>% 
  mutate(date= date(date), year = year(date), bi_pct = central_bank/tot_ownership, con_pct = con_bank/tot_ownership, islam_pct = islam_bank/tot_ownership, ind_pct = individual/tot_ownership, other_pct = others/tot_ownership, mutual_pct = mutual_fund/tot_ownership, ins_pct = ins_pension/tot_ownership, foreign_pct = foreign/tot_ownership) %>% 
   filter(year != 2015) %>% 
  select(date, foreign_pct, bi_pct, con_pct, islam_pct, ind_pct, other_pct, mutual_pct, ins_pct) %>%
  gather(key = "variable", value = "value", -date) %>% 
  ggplot(aes(x = date, y = value))+
  geom_area(aes(fill = variable), alpha = 0.7)+
  theme_classic()

####Others = securities, corporations, and foundations.

# spread pol-rate vs yield
# 2020 spread range between policy rate and yield is quite dispersed
# 2021 spread range is the narrowest
# 
yield_factors %>% 
  na.omit() %>% 
  mutate(year = year(date), spread = abs(domestic_10y-rate)) %>%
  filter(year > 2015) %>% 
  select(date, year, spread, domestic_10y, rate) %>% 
  ggplot(aes(y = spread))+
  geom_boxplot()+
  facet_wrap(~year)


```

From plot above, the movement of domestic 10y yield is parallel with movement of policy rate. The median spread in 2016 until 2018 is quite the same (around 2.5) while median spread of 2019 is the lowest. In 2020 and 2021, the differences between 10y yield and policy rate are increase with median spread is about 3. The range of spread in 2020 is also the widest compared to other observed years.

